---
title: "Lab5_nch"
author: "Nathan Haywood"
date: "2/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
surveys <- read_csv("data/portal_data_joined.csv")
```

## Challenge 1: Using pipes, subset the survey data to include animals collected before 1995 and retain columns year, sex, and weight.
```{r}
subset1 <- surveys %>% 
  filter(year < 1995) %>% 
  select(year, sex, weight)
subset1
```

## Challenge 2: Create a new dataframe with the following criteria: only species ID, and a new column called hindfoot_half where there are no 1/2 hindfood values that are NA or less than 30
```{r}
subset2 <- surveys %>% 
  filter(!is.na(hindfoot_length)) %>% 
  mutate(hindfoot_half= hindfoot_length / 2) %>% 
  filter(hindfoot_half < 30) %>% 
  select(species_id, hindfoot_half)
subset2
```

## Challenge 3a: How many animals were caught in each plot_type
```{r}
count1 = surveys %>% 
  count(plot_type)
count1
```

## Challenge 3b: Use group_by() and summarize() to find the mean, min and max hindfoot lenght for each species
```{r}
subset3 <- surveys %>% 
  filter(!is.na(hindfoot_length)) %>%
  group_by(species_id) %>%
  summarize(min_lenght = min(hindfoot_length), mean_lenght = mean(hindfoot_length), max_lenght = max(hindfoot_length), n = n())
subset3
```

## Challenge 3c: What was the heaviest animal each year? Return year, genus, species_id and weight.
```{r}
subset4 <- surveys %>% 
  filter(!is.na(weight)) %>%
  group_by(year) %>% 
  filter(weight == max(weight)) %>% 
  select(year, genus, species_id, weight)
subset4
```

## Challenge 4a: Spread the surveys data frame with year as columns, plot_id as rows, and the number of genera per plot as the values. You will need to summarize before reshaping, and use the function n_distinct() to get the number of unique genera within a particular chunk of data. It’s a powerful function!
```{r}
subset5 <- surveys %>%
  group_by(plot_id, year) %>%
  summarize(n_genera = n_distinct(genus)) %>%
  spread(year, n_genera)
head(subset5)
```

## Challenge 4b: Now take that data frame and gather() it again, so each row is a unique plot_id by year combination.
```{r}
subset5 %>%
  gather(year, n_genera, plot_id)
```

## Challenge 4c: The surveys data set has two measurement columns: hindfoot_length and weight. This makes it difficult to do things like look at the relationship between mean values of each measurement per year in different plot types. Let’s walk through a common solution for this type of problem. First, use gather() to create a dataset where we have a key column called measurement and a value column that takes on the value of either hindfoot_length or weight.
```{r}
subset6 <- surveys %>%
  gather(measurement, value, hindfoot_length, weight)
subset6
```

## Challenge 4d: With this new data set, calculate the average of each measurement in each year for each different plot_type. Then spread() them into a data set with a column for hindfoot_length and weight.
```{r}
subset6 %>%
  group_by(year, measurement, plot_type) %>%
  summarize(mean_value = mean(value, na.rm = TRUE)) %>%
  spread(measurement, mean_value)
```
